---
title: "week 1 quiz"
author: "Haixu Leng"
date: "5/20/2021"
output:
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.   How many individuals in the Melanoma dataset from the MASS package died from a melanoma?

```{r mass}
library(MASS)
str(Melanoma)
```
Based on the documentation of **MASS** library, the patient who died from melanoma will have status as 1. 
```{r}
n = length(Melanoma$status[Melanoma$status == 1])
```
There are `r n` patients who died from a melanoma.

## 2. What is the average age of individuals in the Melanoma dataset from the MASS package who are alive?

Based on the documentation of **MASS** library, the patient who are alive will have status as 2. 

```{r}
n = length(Melanoma$status[Melanoma$status == 2])
ave_age = mean(Melanoma$age[Melanoma$status == 2])
```


There are `r n` patients who are alive, and the average age is `r ave_age`.

## 3. Which animal in the mammals dataset from the MASS package has the largest brain weight relative to its body weight (that is, the largest brain weight to body weight ratio)?

```{r}
library(MASS)
brain_body_ratio = mammals$brain / mammals$body
row.names(mammals)[which.max(brain_body_ratio)]
```
The answer should be `r row.names(mammals)[which.max(brain_body_ratio)]`.

## 4. Based on this plot, which variable is the most variable? Calculate the standard deviation of this variable. (Use the code block above.)

```{r}
?iris
boxplot(iris)
```

The standard deviation of Petal.Length is `r sd(iris$Petal.Length)`

## 5. 
     str(z)

     min(z[[1]])

     max(z[[2]])

     mean(z[[3]])

## 6. Where were the measurements taken in the airquality dataset?  
```{r}
str(airquality)
?airquality
```
## 7. Using the airquality dataset, what is the average wind speed in May ?

```{r}
wind_in_may = airquality$Wind[airquality$Month == 5]
ave_wind_in_may = mean(wind_in_may)
```

The answer is `r ave_wind_in_may`

## 8. Using the airquality dataset, what is the average ozone measurement? Hint: read the documentation of any function that returns an unexpected result. You will likely find a solution to the issue. 

```{r}
mean(airquality$Ozone, na.rm = TRUE)
```
The answer is `r mean(airquality$Ozone, na.rm = TRUE)`

## 9. Using the airquality dataset, create a scatter plot to compare windspeed and temperature. Based on this plot, you believe that:  

```{r}
plot(airquality$Wind, airquality$Temp)
```
There seems to be a weak correlation between wind speed and temperature. When the temperature increases, the wind speed decreases.

## 10. What proportion of the elements of x are larger than 2 in magnitude?

```{r}
set.seed(1337)
x = rnorm(10000)
index_larger_than_2 = x > 2 | x < -2
num_x_mag_larg_2 = length(index_larger_than_2[index_larger_than_2 == TRUE])
num_x_mag_larg_2 / length(x)
```

The answer is `r num_x_mag_larg_2 / length(x)`

## 11. Write a function called f that has a single argument input with a default value of 42 which is assumed to be a vector of numeric values. The function should output a vector that is input but with any negative values replaced with 0.

```{r}
f <- function(input = c(42)) {
  output = rep(0, length(input)) # initialize output
  for(index in 1:length(input)){
    output[index] = ifelse(input[index] < 0, 0, input[index])
  }
  return(output)
}

# a test
f(c(-1, 2, -3, 4, -5))

# quiz
set.seed(42)
x = rnorm(100, mean = 0, sd = 10)
mean(f(input = x)) - f()
```

## 12. Create three vectors x0, x1, and y. Each should have a length of 30 and store the following:

```{r}
# create vectors x0 and x1
n = 30

# x0: Each element should be the value 1 
x0 = rep(1, n)

#x1: The first 30 square numbers, starting from 1 (so 1, 4, 9, etc.) 
x1 = (1:n)^2

set.seed(42)
y  = 5 * x0 + x1 + rnorm(n = 30, mean = 0 , sd = 1)

```
The mean of y is `r mean(y)`.

## 13. Create a matrix X with columns x0 and x1. Report the sum of the elements in rows 17 and 19.  

```{r}
# x0: Each element should be the value 1 
x0 = rep(1, n)

#x1: The first 30 square numbers, starting from 1 (so 1, 4, 9, etc.) 
x1 = (1:n)^2

m = cbind(x0, x1)
m[17,]
m[19,]
```
The sum is `r sum(m[17,]) + sum(m[19,])`

## 14. Use matrix operations to create a new matrix beta_hat defined as follows. Report the sum of the values stored in this matrix.

$\hat\beta = (X^TX)^{-1}X^Ty$

```{r}
# create vectors x0 and x1
n = 30

# x0: Each element should be the value 1 
x0 = rep(1, n)

#x1: The first 30 square numbers, starting from 1 (so 1, 4, 9, etc.) 
x1 = (1:n)^2

set.seed(42)
y  = 5 * x0 + x1 + rnorm(n = 30, mean = 0 , sd = 1)

x = cbind(x0, x1)

x_t = t(x)

beta = solve(x_t %*% x) %*% (x_t %*%y)
```

The sum of $\hat\beta$ is `r sum(beta)`

## 15. Perform and report the result of the following operation

$\hat y = X\hat\beta$

$\sum_{i=1}^{30}(y_i - \hat y_i)^2$

```{r}
y_hat = x %*% beta
result = sum((y - y_hat)^2)
```

The result is `r result`
