---
title: "Week 9 - Homework"
author: "STAT 420, Summer 2021, haixul2"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---


***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
```

## Exercise 1 (`longley` Macroeconomic Data)

The built-in dataset `longley` contains macroeconomic data for predicting employment. We will attempt to model the `Employed` variable.

```{r, eval = FALSE}
View(longley)
?longley
```

**(a)** What is the largest correlation between any pair of predictors in the dataset?

**Answer**

```{r}
#pairs(longley)
A = cor(longley)
round(A, 3)
max(A[row(A) != col(A)])
```

The maximum correlation between a pair of different predictors is `r max(A[row(A) != col(A)])`. The two predictors are `GNP` and `Year`.


**(b)** Fit a model with `Employed` as the response and the remaining variables as predictors. Calculate and report the variance inflation factor (VIF) for each of the predictors. Which variable has the largest VIF? Do any of the VIFs suggest multicollinearity?

**Answer**

```{r}
model = lm(Employed ~ ., data = longley)
library(faraway)
vif(model)
```
Variable `GNP` has the largest VIF. 

In practice it is common to say that any VIF greater than 5 is cause for concern. So in this example we see there is a huge multicollinearity issue as almost all of the predictors have a VIF greater than 5, the only variable has a VIF **less than** 5 is `Armed.Forces`.


**(c)** What proportion of the observed variation in `Population` is explained by a linear relationship with the other predictors?

**Answer**

```{r}
pop_model = lm(Population ~ . - Employed, data = longley)
summary(pop_model)$r.squared
```

`r summary(pop_model)$r.squared` of the observed variation in `Population` is explained by a linear relationship with the other predictors.

**(d)** Calculate the partial correlation coefficient for `Population` and `Employed` **with the effects of the other predictors removed**.

**Answer**

```{r}
emp_model = lm(Employed ~ . - Population, data = longley)
cor(resid(pop_model), resid(emp_model))
```


**(e)** Fit a new model with `Employed` as the response and the predictors from the model in **(b)** that were significant. (Use $\alpha = 0.05$.) Calculate and report the variance inflation factor for each of the predictors. Which variable has the largest VIF? Do any of the VIFs suggest multicollinearity?

**Answer**

```{r}
summary(model)$coefficients
```

Based on a criteria of $\alpha = 0.05$, the significant predictors are:

  -`Unemployed`
  
  -`Armed.Forces`
  
  -`Year`

```{r}
model_e = lm(Employed ~ Unemployed + Armed.Forces + Year, data = longley)
vif(model_e)
```

**No** multicollinearity issue is found since all VIF values are less than 5.

**(f)** Use an $F$-test to compare the models in parts **(b)** and **(e)**. Report the following:

- The null hypothesis
- The test statistic
- The distribution of the test statistic under the null hypothesis
- The p-value
- A decision
- Which model you prefer, **(b)** or **(e)**

**Answer**

```{r}
anova(model_e, model)
```


The null hypothesis: $\beta$ for `GNP.deflator`, `GNP` and `Population` are zero.

The test statistic: 1.75

The distribution of the test statistic under the null hypothesis: F distribution with degree of freedom as 3

The p-value: 0.23

A decision: fail to reject the null hypothesis

Which model: **(e)**

**(g)** Check the assumptions of the model chosen in part **(f)**. Do any assumptions appear to be violated?

```{r, echo = FALSE}
plot_fitted_resid = function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  plot(fitted(model), resid(model), 
       col = pointcol, pch = 20, cex = 1.5,
       xlab = "Fitted", ylab = "Residuals")
  abline(h = 0, col = linecol, lwd = 2)
}

plot_qq = function(model, pointcol = "dodgerblue", linecol = "darkorange") {
  qqnorm(resid(model), col = pointcol, pch = 20, cex = 1.5)
  qqline(resid(model), col = linecol, lwd = 2)
}
```


**Answer**

```{r}
plot_fitted_resid(model_e)
```

It is **hard to tell** if there is any violation of the equal variance assumpition on the Residual vs Fitted plot, because we only have a small amount of data points. There are two obvious outliers, which are **suspicious**.

```{r}
plot_qq(model_e)
```


From the qqplot, we find there **might** be violations of normality at both ends of the quantiles.

***

## Exercise 2 (`Credit` Data)

For this exercise, use the `Credit` data from the `ISLR` package. Use the following code to remove the `ID` variable which is not useful for modeling.

```{r}
library(ISLR)
data(Credit)
Credit = subset(Credit, select = -c(ID))
```

Use `?Credit` to learn about this dataset.

**(a)** Find a "good" model for `balance` using the available predictors. Use any methods seen in class except transformations of the response. The model should:

- Reach a LOOCV-RMSE below `140`
- Obtain an adjusted $R^2$ above `0.90`
- Fail to reject the Breusch-Pagan test with an $\alpha$ of $0.01$
- Use fewer than 10 $\beta$ parameters

Store your model in a variable called `mod_a`. Run the two given chunks to verify your model meets the requested criteria. If you cannot find a model that meets all criteria, partial credit will be given for meeting at least some of the criteria.

```{r, message = FALSE, warning = FALSE}
library(lmtest)

get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}
```

```{r, eval = FALSE}
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```

**Answer**

```{r, eval = FALSE}
pairs(Credit)
# meets 3 criteria
mod_a0 = lm(Balance ~ (log(Income) + I(log(Rating)^2) + log(Rating) + Limit  + Education + Married + Ethnicity)^2, data = Credit)
mod_a = step(mod_a0, direction = "backward", trace = 0, k = log(length(resid(mod_a0))))

plot_fitted_resid(mod_a)
summary(mod_a)


# test
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```

The original model that I come up with only satisfies 3 out of 4 criteria listed above. but I learned that `Income`, `Limit`, `Rating` and `Student` seem to be significant. But with all these predictors, I cannot pass the `bptest`.

```{r}
pairs(Credit)
# meets 3 criteria
mod_a0 = lm(Balance ~ (log(Income) + sqrt(Limit)  + Education + Married + Ethnicity + Student)^2, data = Credit)
mod_a = step(mod_a0, direction = "backward", trace = 0, k = log(length(resid(mod_a0))))

# After several trials, and some hints from Justin and Alex. I got the following model
mod_a = lm(Balance ~ log(Income) + Limit + Cards + Age + Education + Gender + Student + Married, data = Credit)
plot_fitted_resid(mod_a)
summary(mod_a)


# test
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```

After trying many new models and with some help from Justin and Alex, I get a model that **satisfies all the criteria**.


**(b)** Find another "good" model for `balance` using the available predictors. Use any methods seen in class except transformations of the response. The model should:

- Reach a LOOCV-RMSE below `130`
- Obtain an adjusted $R^2$ above `0.85`
- Fail to reject the Shapiro-Wilk test with an $\alpha$ of $0.01$
- Use fewer than 25 $\beta$ parameters

Store your model in a variable called `mod_b`. Run the two given chunks to verify your model meets the requested criteria. If you cannot find a model that meets all criteria, partial credit will be given for meeting at least some of the criteria.

```{r, message = FALSE, warning = FALSE}
library(lmtest)

get_bp_decision = function(model, alpha) {
  decide = unname(bptest(model)$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_sw_decision = function(model, alpha) {
  decide = unname(shapiro.test(resid(model))$p.value < alpha)
  ifelse(decide, "Reject", "Fail to Reject")
}

get_num_params = function(model) {
  length(coef(model))
}

get_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

get_adj_r2 = function(model) {
  summary(model)$adj.r.squared
}
```

```{r, eval = FALSE}
get_loocv_rmse(mod_b)
get_adj_r2(mod_b)
get_sw_decision(mod_b, alpha = 0.01)
get_num_params(mod_b)
```

**Answer**


```{r}
#mod_a = lm(Balance~Limit + log(Income) + Student + Cards + Limit:log(Income) + 
#Limit:Student + log(Income):Student + Limit:Cards + Limit:log(Income):Student, data = Credit)
mod_a0 = lm(Balance ~ (log(Income) + Limit + Cards + Age + Education + Gender + Student + Married)^3, data = Credit)
mod_a = lm(Balance ~ log(Income) + Limit + Cards + Age + Education + Gender + Student + Married + Limit:Age + log(Income):Education, data = Credit)

summary(mod_a)
#plot_fitted_resid(mod_a)
# test
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```

The model that I come up with **almost satisfies** all criteria listed above. The only issue is that my LOOCV RMSE is 130.2, which is slightly higher than 130.

I find that it is hard to fail to reject the null hypothesis with `Student` as a predictor. However, without `Student`, the LOOCV-RMSE cannot be lowered to a small value. I am not able to find a model to achieve that. The core of this issue is that there are a lot of data points with a balance of 0, and they cannot be fitted well. **If I remove these 0 value balance points**, the model can perform much better (see below).

```{r}
# meets 3 criteria
mod_a0 = lm(Balance ~ (log(Income) + I(log(Rating)^2) + log(Rating) + Limit + Student + Education + Married + Ethnicity)^2, data = Credit[Credit$Balance>0, ])
mod_a = step(mod_a0, direction = "backward", trace = 0, k = log(length(resid(mod_a0))))
plot(log(Credit$Income), Credit$Balance,
     xlab = "log(Income)",
     ylab = "Balance",
     main = "Unusual points near balance = 0")

plot_fitted_resid(mod_a)
summary(mod_a)


# test
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
get_bp_decision(mod_a, alpha = 0.01)
get_num_params(mod_a)
```


***

## Exercise 3 (`Sacramento` Housing Data)

For this exercise, use the `Sacramento` data from the `caret` package. Use the following code to perform some preprocessing of the data.

```{r}
library(caret)
library(ggplot2)
data(Sacramento)
sac_data = Sacramento
sac_data$limits = factor(ifelse(sac_data$city == "SACRAMENTO", "in", "out"))
sac_data = subset(sac_data, select = -c(city, zip))
```

Instead of using the `city` or `zip` variables that exist in the dataset, we will simply create a variable (`limits`) indicating whether or not a house is technically within the city limits of Sacramento. (We do this because they would both be factor variables with a **large** number of levels. This is a choice that is made due to laziness, not necessarily because it is justified. Think about what issues these variables might cause.)

Use `?Sacramento` to learn more about this dataset.

A plot of longitude versus latitude gives us a sense of where the city limits are.

```{r}
qplot(y = longitude, x = latitude, data = sac_data,
      col = limits, main = "Sacramento City Limits ")
```

After these modifications, we test-train split the data.

```{r}
set.seed(420)
sac_trn_idx  = sample(nrow(sac_data), size = trunc(0.80 * nrow(sac_data)))
sac_trn_data = sac_data[sac_trn_idx, ]
sac_tst_data = sac_data[-sac_trn_idx, ]
```

The training data should be used for all model fitting. Our goal is to find a model that is useful for predicting home prices.

**(a)** Find a "good" model for `price`. Use any methods seen in class. The model should reach a LOOCV-RMSE below 77,500 in the training data. Do not use any transformations of the response variable.

**Answer**

```{r}
plot(sac_trn_data)
mod_a0 = lm(price ~ . + log(beds), data = sac_trn_data)
mod_a = step(mod_a0, direction = "backward", trace = 0, k = log(length(resid(mod_a0))))
summary(mod_a)


# test
get_loocv_rmse(mod_a)
get_adj_r2(mod_a)
```

The model I find has a LOOCV-RMSE **below 77500.**


**(b)** Is a model that achieves a LOOCV-RMSE below 77,500 useful in this case? That is, is an average error of 77,500 low enough when predicting home prices? To further investigate, use the held-out test data and your model from part **(a)** to do two things:

- Calculate the average percent error:
\[
\frac{1}{n}\sum_i\frac{|\text{predicted}_i - \text{actual}_i|}{\text{predicted}_i} \times 100
\]
- Plot the predicted versus the actual values and add the line $y = x$.

Based on all of this information, argue whether or not this model is useful.

**Answer** 

```{r}
test_price = sac_tst_data$price
predict_price = predict(mod_a, sac_tst_data)

ave_per_error = mean(abs(test_price - predict_price) / predict_price) * 100
plot(predict_price, test_price)
abline(a = 0, b = 1)

ave_per_error
```

  - The average percent error: `r ave_per_error`
  
  - Plot is displayed above
  
I think the model is still useful, since the predicted price vs test price is close to `y = x`. There are several things that look suspicious from the plot:

  - The variation at low price houses seems to be smaller than the variation of more expensive houses
  
  - There is an outlier priced around 800k, but my model suggests it should be around 250k 

***

## Exercise 4 (Does It Work?)

In this exercise, we will investigate how well backwards AIC and BIC actually perform. For either to be "working" correctly, they should result in a low number of both **false positives** and **false negatives**. In model selection,

- **False Positive**, FP: Incorrectly including a variable in the model. Including a *non-significant* variable
- **False Negative**, FN: Incorrectly excluding a variable in the model. Excluding a *significant* variable

Consider the **true** model

\[
Y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + \beta_4 x_4 + \beta_5 x_5 + \beta_6 x_6 + \beta_7 x_7 + \beta_8 x_8 + \beta_9 x_9 + \beta_{10} x_{10} + \epsilon
\]

where $\epsilon \sim N(0, \sigma^2 = 4)$. The true values of the $\beta$ parameters are given in the `R` code below.

```{r}
beta_0  = 1
beta_1  = -1
beta_2  = 2
beta_3  = -2
beta_4  = 1
beta_5  = 1
beta_6  = 0
beta_7  = 0
beta_8  = 0
beta_9  = 0
beta_10 = 0
sigma = 2
```

Then, as we have specified them, some variables are significant, and some are not. We store their names in `R` variables for use later.

```{r}
not_sig  = c("x_6", "x_7", "x_8", "x_9", "x_10")
signif = c("x_1", "x_2", "x_3", "x_4", "x_5")
```

We now simulate values for these `x` variables, which we will use throughout part **(a)**.

```{r}
set.seed(420)
n = 100
x_1  = runif(n, 0, 10)
x_2  = runif(n, 0, 10)
x_3  = runif(n, 0, 10)
x_4  = runif(n, 0, 10)
x_5  = runif(n, 0, 10)
x_6  = runif(n, 0, 10)
x_7  = runif(n, 0, 10)
x_8  = runif(n, 0, 10)
x_9  = runif(n, 0, 10)
x_10 = runif(n, 0, 10)
```

We then combine these into a data frame and simulate `y` according to the true model.

```{r}
sim_data_1 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma)
)
```

We do a quick check to make sure everything looks correct.

```{r}
head(sim_data_1)
```

Now, we fit an incorrect model.

```{r}
fit = lm(y ~ x_1 + x_2 + x_6 + x_7, data = sim_data_1)
coef(fit)
```

Notice, we have coefficients for `x_1`, `x_2`, `x_6`, and `x_7`. This means that `x_6` and `x_7` are false positives, while `x_3`, `x_4`, and `x_5` are false negatives.

To detect the false negatives, use:

```{r}
# which are false negatives?
!(signif %in% names(coef(fit)))
```

To detect the false positives, use:

```{r}
# which are false positives?
names(coef(fit)) %in% not_sig
```

Note that in both cases, you could `sum()` the result to obtain the number of false negatives or positives.

**(a)** Set a seed equal to your birthday; then, using the given data for each `x` variable above in `sim_data_1`, simulate the response variable `y` 300 times. Each time,

- Fit an additive model using each of the `x` variables.
- Perform variable selection using backwards AIC.
- Perform variable selection using backwards BIC.
- Calculate and store the number of false negatives for the models chosen by AIC and BIC.
- Calculate and store the number of false positives for the models chosen by AIC and BIC.

Calculate the rate of false positives and negatives for both AIC and BIC. Compare the rates between the two methods. Arrange your results in a well formatted table.

**Answer**

```{r}
n_sim = 300
set.seed(19900210)
AIC_false_negative = rep(0, n_sim)
AIC_false_positive = rep(0, n_sim)
BIC_false_negative = rep(0, n_sim)
BIC_false_positive = rep(0, n_sim)

for(i in 1:n_sim){
  sim_data_1 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma))
  
  # fit an additive model
  fit = lm(y ~ ., data = sim_data_1)
  # AIC
  fit_aic = step(fit, direction = "backward", trace = 0)
  # BIC
  fit_bic = step(fit, direction = "backward", trace = 0, k = log(length(resid(fit))))
  # Store the results
  AIC_false_negative[i] = sum(!(signif %in% names(coef(fit_aic))))
  BIC_false_negative[i] = sum(!(signif %in% names(coef(fit_bic))))
  AIC_false_positive[i] = sum(names(coef(fit_aic)) %in% not_sig)
  BIC_false_positive[i] = sum(names(coef(fit_bic)) %in% not_sig)
}
```


```{r}
table = data.frame(c("# of false negative", "# of false positive", "FN / 300", "FP / 300"), 
                   c(sum(AIC_false_negative), sum(AIC_false_positive), sum(AIC_false_negative) / 300, sum(AIC_false_positive) / 300),
                   c(sum(BIC_false_negative), sum(BIC_false_positive), 
                     sum(BIC_false_negative) / 300, sum(BIC_false_positive) / 300))
names(table) = c("", "AIC", "BIC" )
knitr::kable(table)
```

AIC method seems to give higher false positive rate than BIC.

**(b)** Set a seed equal to your birthday; then, using the given data for each `x` variable below in `sim_data_2`, simulate the response variable `y` 300 times. Each time,

- Fit an additive model using each of the `x` variables.
- Perform variable selection using backwards AIC.
- Perform variable selection using backwards BIC.
- Calculate and store the number of false negatives for the models chosen by AIC and BIC.
- Calculate and store the number of false positives for the models chosen by AIC and BIC.

Calculate the rate of false positives and negatives for both AIC and BIC. Compare the rates between the two methods. Arrange your results in a well formatted table. Also compare to your answers in part **(a)** and suggest a reason for any differences.

```{r}
set.seed(94)
x_1  = runif(n, 0, 10)
x_2  = runif(n, 0, 10)
x_3  = runif(n, 0, 10)
x_4  = runif(n, 0, 10)
x_5  = runif(n, 0, 10)
x_6  = runif(n, 0, 10)
x_7  = runif(n, 0, 10)
x_8  = x_1 + rnorm(n, 0, 0.1)
x_9  = x_1 + rnorm(n, 0, 0.1)
x_10 = x_2 + rnorm(n, 0, 0.1)

sim_data_2 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma)
)
```


**Answer**

```{r}
n_sim = 300
set.seed(19900210)
AIC_false_negative = rep(0, n_sim)
AIC_false_positive = rep(0, n_sim)
BIC_false_negative = rep(0, n_sim)
BIC_false_positive = rep(0, n_sim)

for(i in 1:n_sim){
  sim_data_1 = data.frame(x_1, x_2, x_3, x_4, x_5, x_6, x_7, x_8, x_9, x_10,
  y = beta_0 + beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + 
      beta_5 * x_5 + rnorm(n, 0 , sigma))
  
  # fit an additive model
  fit = lm(y ~ ., data = sim_data_1)
  # AIC
  fit_aic = step(fit, direction = "backward", trace = 0)
  # BIC
  fit_bic = step(fit, direction = "backward", trace = 0, k = log(length(resid(fit))))
  # Store the results
  AIC_false_negative[i] = sum(!(signif %in% names(coef(fit_aic))))
  BIC_false_negative[i] = sum(!(signif %in% names(coef(fit_bic))))
  AIC_false_positive[i] = sum(names(coef(fit_aic)) %in% not_sig)
  BIC_false_positive[i] = sum(names(coef(fit_bic)) %in% not_sig)
}

# table
table = data.frame(c("# of false negative", "# of false positive", "FN / 300", "FP / 300"), 
                   c(sum(AIC_false_negative), sum(AIC_false_positive), sum(AIC_false_negative) / 300, sum(AIC_false_positive) / 300),
                   c(sum(BIC_false_negative), sum(BIC_false_positive), 
                     sum(BIC_false_negative) / 300, sum(BIC_false_positive) / 300))
names(table) = c("", "AIC", "BIC" )
knitr::kable(table)
```

Comparing to (a) both methods give a higher false positive rate and a higher false negative rate.

I think the reason behind the change is **colinearity**. For example, `x_8`, `x_9`, and `x_1` are basically the same predictor. If the method picks `x_8` instead of `x_1`, the model has the same prediction power as the model with predictor `x_1` instead of `x_8`. However, the model without `x_1` but with `x_8` will be counted as a false negative and a false positive, because $\beta_1 \neq0,\beta_8=0$. 