---
title: "Week 1 practice quiz"
author: "Haixu Leng"
date: "5/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. $x=1:100$, find the value of $\sum\limits_{i}^n ln(x_{i})$

```{r cars}
x = c(1:100)
y = sum(log(x))

```
The result is equal to `r y`.

## 2. How many of the elements of a_vector are greater than or equal to 5?

     set.seed(42)
     a_vector = rpois(250, lambda = 6)

```{r}
set.seed(42)
a_vector = rpois(250, lambda = 6)
n = length(a_vector[a_vector >= 5])
```
There are `r n` elements in `a_vector` that are greater than or equal to 5.

## 3. Create a new vector **y**, which adds 5 to the elements stored in odd indices of **x** and subtracts 10 from the elements stored in even indices of **x**. Calculate the standard deviation of this new vector.

```{r}
x = 1:100
odd_vector = rep(c(1,0), length(x) / 2)
even_vector = rep(c(0,1), length(x) / 2)
y = x + 5 * odd_vector - 10 * even_vector
```
Result: `r sd(y)`.

## 4. Which of the following options would return the third element of the list quiz_list? (Select all that apply.)
```{r echo=TRUE}
quiz_list = list(
  x = c(1, 2),
  y = "Hello Quiz Taker",
  z = "zz"
)

quiz_list[3]
quiz_list[[3]]
quiz_list["3"] # does not work
quiz_list$z
#quiz_list$3 does not work
```

## 5. Create a histogram of age in the Melanoma dataset from the MASS package. How would you describe this data?
```{r include=FALSE}
library(MASS)
MASS::Melanoma
```
```{r}
hist(Melanoma$age)
hist(rchisq(1500, df = 5), breaks = 20) # example of a right skewed data
```


