---
title: "week_2_practice_quiz"
author: "Haixu Leng"
date: "5/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Consider a random variable X that has a normal distribution with a mean of 5 and a variance of 9.

Calculate $P[X > 4]$.

```{r}
pnorm(4, mean = 5, sd =3, lower.tail = FALSE)
```

## 2. What is the expected value of Y give that x = 5.
$Y = -3 + 2.5x, \epsilon \sim N(0, \sigma^2 = 4)$

```{r}
y = -3 + 2.5 * 5
y
```

## 3. Given the SLR model in Question 2, what is the standard deviation of Y when x is 10?

The standard deviation should be determined by $\sigma$ of $\epsilon$, which is $2$.


## 4. For this question, use the built-in trees dataset in R. Fit a simple linear regression model with Girth as the response and Height as the predictor.

What is the slope of the fitted regression line?

```{r}
result = lm(trees$Girth ~ trees$Height)
```

The slope is `r coefficients(result)[[2]]`

## 5. What is the value of $R^2$ for this fitted SLR model?

```{r}
x = trees$Height
y = trees$Girth
y_hat = coefficients(result)[[1]] + x * coefficients(result)[[2]]
SST = sum((y - mean(y))^2)
SSReg = sum((y_hat - mean(y))^2)
SSE = sum((y_hat - y)^2)
R_2 = SSReg / SST
```

Result: $R^2 =$ `r R_2`.

