---
title: "Week 4 - Homework"
author: "STAT 420, Summer 2021, Haixu Leng (NetID: haixul2)"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


## Exercise 1 (Using `lm`)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` 
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

**(a)** Fit the following multiple linear regression model in `R`. Use `Calories` as the response and `Fat`, `Sugar`, and `Sodium` as predictors.

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \epsilon_i.
\]

Here,

- $Y_i$ is `Calories`.
- $x_{i1}$ is `Fat`.
- $x_{i2}$ is `Sugar`.
- $x_{i3}$ is `Sodium`.

Use an $F$-test to test the significance of the regression. Report the following:
 
- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

**Answer:**

```{r}
nutrition = read.csv("nutrition-2018.csv")
nutrition_model = lm(Calories ~ Fat + Sugar + Sodium, data = nutrition)
```

- Null hypothesis $H_0: \beta_1 = \beta_2=...=\beta_{p-1} = 0$, the alternative hypothesis is $H_1: at \ least \ one \ of \ \beta_j \neq0, j=1,2,..., p-1$ 

- The value is `r summary(nutrition_model)$fstatistic["value"]`.

- The p-value of this test is `r pf(summary(nutrition_model)$fstatistic["value"], summary(nutrition_model)$fstatistic["numdf"], summary(nutrition_model)$fstatistic["dendf"], lower.tail = FALSE)`. It is so small, so 0 is returned.

- We can reject the null hypothesis.

- At least one of the predictors (Fat, Sodium, and Sugar) has a significant linear relationship with the response (Calories).

**(b)** Output only the estimated regression coefficients. Interpret all $\hat{\beta}_j$ coefficients in the context of the problem.

**Answer:**

```{r}
coefficients(nutrition_model)
```

- Intercept: When there is no fat, sodium and sugar, the calories is `r coefficients(nutrition_model)["(Intercept)"]`

- Fat: for an increase in fat by 1 unit, we will get an increase of calories by `r coefficients(nutrition_model)["Fat"]`

- Sodium: for an increase in sodium by 1 unit, we will get an increase of calories by `r coefficients(nutrition_model)["Sodium"]`

- Sugar: for an increase in sugar by 1 unit, we will get an increase of calories by `r coefficients(nutrition_model)["Sugar"]`

**(c)** Use your model to predict the number of `Calories` in a Filet-O-Fish. According to [McDonald's publicized nutrition facts](https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html), the Filet-O-Fish contains 18g of fat, 5g of sugar, and 580mg of sodium.

**Answer:**

```{r}
predict(nutrition_model, newdata = data.frame(Fat = 18, Sugar = 5, Sodium = 580))
```

It is `r predict(nutrition_model, newdata = data.frame(Fat = 18, Sodium = 580, Sugar = 5))`

**(d)** Calculate the standard deviation, $s_y$, for the observed values in the Calories variable. Report the value of $s_e$ from your multiple regression model. Interpret both estimates in the context of this problem.

**Answer:**

```{r}
# calculate the standard deviation using the residuals
s_y = sqrt(sum((nutrition_model$residuals)^2) / (length(nutrition_model$residuals)))

s_e = summary(nutrition_model)$sigma

```

The results of $s_y = `r s_y`$ and $s_e = `r s_e`$ are very close to each other. 

\[
s_y = \sqrt{\frac{\sum (y_i - \hat y_i)^2}{n}}
\]

and 

\[
s_e = \sqrt{\frac{\sum (y_i - \hat y_i)^2}{n - p}}, p = 4
\]

The difference is that foe $s_e$, we have to consider the degree of freedom in the modeling, but in $s_y$ we simply calculate the variance.

**(e)** Report the value of $R^2$ for the model. Interpret its meaning in the context of the problem.

```{r}
summary(nutrition_model)$r.squared
```


**Answer:** $R^2 = `r summary(nutrition_model)["r.squared"]`$. It means that `r summary(nutrition_model)["r.squared"]` proportion of observed variation in y that can be explained by the linear regression model.

**(f)** Calculate a 90% confidence interval for $\beta_2$. Give an interpretation of the interval in the context of the problem.

**Answer:**

```{r}
(beta_2_int = confint(nutrition_model, parm = "Sugar", level = 0.9))
```

We are 90% confident that the mean of $\beta_2$ will be from `r beta_2_int[[1]]` to `r beta_2_int[[2]]`

**(g)** Calculate a 95% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

**Answer:** 

```{r}
(beta_0_int = confint(nutrition_model, parm = "(Intercept)", level = 0.95))
```

We are 95% confident that the mean of $\beta_0$ will be from `r beta_0_int[[1]]` to `r beta_0_int[[2]]`


**(h)** Use a 99% confidence interval to estimate the mean Calorie content of a food with 15g of fat, 0g of sugar, and 260mg of sodium, which is true of a medium order of McDonald's french fries. Interpret the interval in context.

**Answer:**

```{r}
(fry = predict(nutrition_model, newdata = data.frame(Fat = 15, Sugar = 0, Sodium = 260), interval = c("confidence"), level = 0.99))
```

We are 99% confident that the mean calories value of a medium order of McDonald's french fries will be from `r fry[[2]]` to `r fry[[3]]`

 
**(i)** Use a 99% prediction interval to predict the Calorie content of a Crunchy Taco Supreme, which has 11g of fat, 2g of sugar, and 340mg of sodium according to [Taco Bell's publicized nutrition information](https://www.tacobell.com/nutrition/info). Interpret the interval in context.

**Answer:**

```{r}
(taco = predict(nutrition_model, newdata = data.frame(Fat = 11, Sugar = 2, Sodium = 340), interval = c("prediction"), level = 0.99))
```

We are 99% confident that a new observed calories value of a Crunchy Taco Supreme will be from `r taco[[2]]` to `r taco[[3]]` 


***

## Exercise 2 (More `lm` for Multiple Regression)

For this exercise we will use the data stored in [`goalies17.csv`](goalies17.csv). It contains career data for goaltenders in the National Hockey League during the first 100 years of the league from the 1917-1918 season to the 2016-2017 season. It holds the 750 individuals who played at least one game as goalie over this timeframe. The variables in the dataset are:
 
- `Player` - Player's Name (those followed by * are in the Hall of Fame as of 2017)
- `First` - First year with game recorded as goalie
- `Last` - Last year with game recorded as goalie
- `Active` - Number of seasons active in the NHL
- `GP` - Games Played
- `GS` - Games Started
- `W` - Wins
- `L` - Losses (in regulation)
- `TOL` - Ties and Overtime Losses
- `GA` - Goals Against
- `SA` - Shots Against
- `SV` - Saves
- `SV_PCT` - Save Percentage
- `GAA` - Goals Against Average
- `SO` - Shutouts
- `PIM` - Penalties in Minutes
- `MIN` - Minutes

For this exercise we will consider three models, each with Wins as the response. The predictors for these models are:

- Model 1: Goals Against, Saves
- Model 2: Goals Against, Saves, Shots Against, Minutes, Shutouts
- Model 3: All Available

After reading in the data but prior to any modeling, you should clean the data set for this exercise by removing the following variables: `Player`, `GS`, `L`, `TOL`, `SV_PCT`, and `GAA`.

**(a)** Use an $F$-test to compares Models 1 and 2. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

**Answer:**

```{r}
# load data
goalies17 = read.csv("goalies17.csv")

# clean the data set for this exercise by removing the following variables: Player, GS, L, TOL, SV_PCT, and GAA.
goalies17 = subset(goalies17, select = -c(Player, GS, L, TOL, SV_PCT, GAA))

# get models
model_1 = lm(GA ~ SV, data = goalies17)
model_2 = lm(GA ~ SV + SA + MIN + SO, data = goalies17)

# anova test
(test1 = anova(model_1, model_2))
```

- null hypothesis $H_0$: the coefficients for Saves, Shots Against, Minutes, Shutouts are zero, $\beta_2=\beta_3=\beta_4=0$
- the value of the test is `r test1$F[2]`
- the p-value is `r test1["Pr(>F)"][2,1]`
- since the p value is less than $\alpha$, we can reject the null hypothesis
- Model 2 is better, because at least some of the extra parameters are significant

**Note**: Row 25 has NA value in it, for (a) I did not remove this row.


**(b)** Use an $F$-test to compare Model 3 to your preferred model from part **(a)**. Report the following:

- The null hypothesis
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- The model you prefer

**Answer:**

```{r}
# remove row 25, since there is NA in W
goalies17 = goalies17[-c(25), ]

# model 3
model_3 = lm(GA~., data = goalies17)

# redo model 2
model_2 = lm(GA ~ SV + SA + MIN + SO, data = goalies17)

# anova test
(test2 = anova(model_2, model_3))
```

- the null hypothesis $H_0$: the extra $\beta_i$ for parameters not in model_2 are zero.
- The value of the test is `r test2$F[2]`
- the p-value is `r test2["Pr(>F)"][2,1]`
- since the p value is less than $\alpha$, we can reject the null hypothesis
- Model 3 is better, because at least some of the extra parameters are significant

**Note:** row 25 is removed to have consistent data before model_2 and model_3

**(c)** Use a $t$-test to test $H_0: \beta_{\texttt{SV}} = 0 \ \text{vs} \ H_1: \beta_{\texttt{SV}} \neq 0$ for the model you preferred in part **(b)**. Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

**Answer:** 

```{r}
coef(summary(model_3))["SV", ]
```

- The value of the test statistic is `r coef(summary(model_3))["SV", "t value"]`

- The p-value is `r coef(summary(model_3))["SV", "Pr(>|t|)"]`

- Since the p-value is much less than $\alpha$, we reject the null hypothesis

***

## Exercise 3 (Regression without `lm`)

For this exercise we will once again use the `Ozone` data from the `mlbench` package. The goal of this exercise is to fit a model with `ozone` as the response and the remaining variables as predictors.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Obtain the estimated regression coefficients **without** the use of `lm()` or any other built-in functions for regression. That is, you should use only matrix operations. Store the results in a vector `beta_hat_no_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_no_lm ^ 2)`.

**Answer:**

```{r}
# get x and y
y = as.vector(Ozone[,"ozone"])
x = as.matrix(cbind(b0 = rep(1, length(y)), Ozone[,c("wind", "humidity", "temp")]))

(beta_hat_no_lm = as.vector(solve(t(x) %*% x) %*% t(x) %*% y))
sum(beta_hat_no_lm^2)
```

The results are shown above.


**(b)** Obtain the estimated regression coefficients **with** the use of `lm()`. Store the results in a vector `beta_hat_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_lm ^ 2)`.

**Answer:**

```{r}
ozone_model = lm(ozone ~ ., data = Ozone)
(beta_hat_lm = as.vector(coefficients(ozone_model)))
sum(beta_hat_lm^2)
```

Results are listed above, and they are consistent with the no lm results.


**(c)** Use the `all.equal()` function to verify that the results are the same. You may need to remove the names of one of the vectors. The `as.vector()` function will do this as a side effect, or you can directly use `unname()`.

**Answer:** 

```{r}
all.equal(beta_hat_no_lm, beta_hat_lm)
```

It is confirmed that the results are the same.

**(d)** Calculate $s_e$ without the use of `lm()`. That is, continue with your results from **(a)** and perform additional matrix operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

**Answer:**

```{r}
y_hat = x %*% beta_hat_no_lm
# residual
e = y - y_hat
(se = sqrt(t(e) %*% e / (length(y) - 4)))

# verify with lm
summary(ozone_model)$sigma
```

The manually calculated value is the same as the returned value from `lm()`, which is `r summary(ozone_model)$sigma`


**(e)** Calculate $R^2$ without the use of `lm()`. That is, continue with your results from **(a)** and **(d)**, and perform additional operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.


**Answer:**

```{r}
y_m = rep(mean(y), length(y))
e_tot = y - y_m
e_reg = y_hat - y_m

SST = t(e_tot) %*% e_tot
SSReg = t(e_reg) %*% e_reg
# calculated value
SSReg / SST
# returned value from lm
summary(ozone_model)$r.squared
```

Calculated $R^2$ is the same as the returned $R^2$ from `lm()`, which is `r summary(ozone_model)$r.squared`

***

## Exercise 4 (Regression for Prediction)

For this exercise use the `Auto` dataset from the `ISLR` package. Use `?Auto` to learn about the dataset. The goal of this exercise is to find a model that is useful for **predicting** the response `mpg`. We remove the `name` variable as it is not useful for this analysis. (Also, this is an easier to load version of data from the textbook.)

```{r}
# load required package, remove "name" variable
library(ISLR)
Auto = subset(Auto, select = -c(name))
dim(Auto)
```

When evaluating a model for prediction, we often look at RMSE. However, if we both fit the model with all the data as well as evaluate RMSE using all the data, we're essentially cheating. We'd like to use RMSE as a measure of how well the model will predict on *unseen* data. If you haven't already noticed, the way we had been using RMSE resulted in RMSE decreasing as models became larger.

To correct for this, we will only use a portion of the data to fit the model, and then we will use leftover data to evaluate the model. We will call these datasets **train** (for fitting) and **test** (for evaluating). The definition of RMSE will stay the same

\[
\text{RMSE}(\text{model, data}) = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}
\]

where

- $y_i$ are the actual values of the response for the given data.
- $\hat{y}_i$ are the predicted values using the fitted model and the predictors from the data.

However, we will now evaluate it on both the **train** set and the **test** set separately. So each model you fit will have a **train** RMSE and a **test** RMSE. When calculating **test** RMSE, the predicted values will be found by predicting the response using the **test** data with the model fit using the **train** data. *__Test__ data should never be used to fit a model.*

- Train RMSE: Model fit with *train* data. Evaluate on **train** data.
- Test RMSE: Model fit with *train* data. Evaluate on **test** data.

Set a seed of `22`, and then split the `Auto` data into two datasets, one called `auto_trn` and one called `auto_tst`. The `auto_trn` data frame should contain 290 randomly chosen observations. The `auto_tst` data will contain the remaining observations. Hint: consider the following code:

```{r, eval = FALSE}
set.seed(22)
auto_trn_idx = sample(1:nrow(Auto), 290)
```

Fit a total of five models using the training data.

- One must use all possible predictors.
- One must use only `displacement` as a predictor.
- The remaining three you can pick to be anything you like. One of these should be the *best* of the five for predicting the response.

For each model report the **train** and **test** RMSE. Arrange your results in a well-formatted markdown table. Argue that one of your models is the best for predicting the response.

**Answer:**

```{r}
set.seed(22)
auto_trn_idx = sample(1:nrow(Auto), 290)

# get the train and test data
train = Auto[auto_trn_idx, ]
test = Auto[-auto_trn_idx, ]

# define a function for rmse
rmse = function(model, dataset){
  residuals = dataset$mpg - predict(model, newdata = dataset)
  sqrt(sum((residuals)^2) / length(residuals))
}

# model 1
model_1 = lm(mpg ~ ., data = train)

# model 2
model_2 = lm(mpg ~ displacement, data = train)

# model 3
model_3 = lm(mpg ~ displacement + weight + year + origin + cylinders + acceleration, data = train)

# model 4
model_4 = lm(mpg ~ displacement + weight + year + origin + cylinders, data = train)

# model 5
model_5 = lm(mpg ~ displacement + weight + year + origin, data = train)

models = data.frame(c("model 1", "model 2", "model 3", "model 4", "model 5"), rep(0,5), rep(0,5), rep(0,5), rep(0,5), rep(0,5), rep(0,5), rep(0,5), rep(0,5), rep(0,5))
colnames(models) = c("", "cylinders", "displacement", "horsepower", "weight", "year", "origin", "acceleration", "rmse_train", "rmse_test")

models[1, 2:8] = 1
models[1, 9] = rmse(model_1, train)
models[1, 10] = rmse(model_1, test)

models[2, 3] = 1
models[2, 9] = rmse(model_2, train)
models[2, 10] = rmse(model_2, test)

models[3, 2:8] = 1
models[3, 4] = 0
models[3, 9] = rmse(model_3, train)
models[3, 10] = rmse(model_3, test)

models[4, 2:8] = 1
models[4, 4] = 0
models[4, 8] = 0
models[4, 9] = rmse(model_4, train)
models[4, 10] = rmse(model_4, test)

models[5, 2:8] = 1
models[5, 4] = 0
models[5, 8] = 0
models[5, 2] = 0
models[5, 9] = rmse(model_5, train)
models[5, 10] = rmse(model_5, test)

knitr::kable(models)
```

**Model 4** is the best model among the 5 models, even though its $RMSE_{train}$ is slightly greater than model 1, but it has the least $RMSE_{test}$. This suggests that model 4 does well in both the train group and the test group. In model 4, I used "displacement + weight + year + origin + cylinders" as predictors.

***

## Exercise 5 (Simulating Multiple Regression)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \beta_4 x_{i4} + \beta_5 x_{i5} + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 2$
- $\beta_1 = -0.75$
- $\beta_2 = 1.6$
- $\beta_3 = 0$
- $\beta_4 = 0$
- $\beta_5 = 2$
- $\sigma^2 = 25$

We will use samples of size `n = 40`.

We will verify the distribution of $\hat{\beta}_1$ as well as investigate some hypothesis tests.

**(a)** We will first generate the $X$ matrix and data frame that will be used throughout the exercise. Create the following nine variables:

- `x0`: a vector of length `n` that contains all `1`
- `x1`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `x2`: a vector of length `n` that is randomly drawn from a uniform distribution between `0` and `4`
- `x3`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `1`
- `x4`: a vector of length `n` that is randomly drawn from a uniform distribution between `-2` and `2`
- `x5`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `X`: a matrix that contains `x0`, `x1`, `x2`, `x3`, `x4`, and `x5` as its columns
- `C`: the $C$ matrix that is defined as $(X^\top X)^{-1}$
- `y`: a vector of length `n` that contains all `0`
- `sim_data`: a data frame that stores `y` and the **five** *predictor* variables. `y` is currently a placeholder that we will update during the simulation.

Report the sum of the diagonal of `C` as well as the 5th row of `sim_data`. For this exercise we will use the seed `420`. Generate the above variables in the order listed after running the code below to set a seed.

```{r}
set.seed(400)
sample_size = 40
```

**Answer:**

```{r}
x0 = rep(1, sample_size)
x1 = rnorm(sample_size, mean = 0, sd = 2)
x2 = runif(sample_size, min = 0, max = 4)
x3 = rnorm(sample_size, mean = 0, sd = 1)
x4 = runif(sample_size, min = -2, max = 2)
x5 = rnorm(sample_size, mean = 0, sd = 2)

X = as.matrix(cbind(x0, x1, x2, x3, x4, x5))

C = solve(t(X) %*% X)

y = rep(0, sample_size)

sim_data = data.frame(y = y, x1 = x1, x2 = x2, x3 = x3, x4 = x4, x5 = x5)

sum(diag(C))

sum(sim_data[5,])
```

The sum of diagonal of C is `r sum(diag(C))`, and the sum of the 5th row of `sim_data` is `r sum(sim_data[5,])`.

**(b)** Create three vectors of length `2500` that will store results from the simulation in part **(c)**. Call them `beta_hat_1`, `beta_3_pval`, and `beta_5_pval`.

**Answer:**

```{r}
beta_hat_1 = rep(0, 2500)
beta_3_pval = rep(0, 2500)
beta_5_pval = rep(0, 2500)
```


**(c)** Simulate 2500 samples of size `n = 40` from the model above. Each time update the `y` value of `sim_data`. Then use `lm()` to fit a multiple regression model. Each time store:

- The value of $\hat{\beta}_1$ in `beta_hat_1`
- The p-value for the two-sided test of $\beta_3 = 0$ in `beta_3_pval`
- The p-value for the two-sided test of $\beta_5 = 0$ in `beta_5_pval`

**Answer:**

```{r}
sim = function(){
  beta = c(2, -0.75, 1.6, 0, 0, 2)
  return(X %*% beta + rnorm(40, 0, 5))
}

for(i in 1:2500){
  y_sim = sim()
  sim_data["y"] = y_sim
  sim_model = lm(y ~ ., data = sim_data)
  beta_hat_1[i] = coef(sim_model)["x1"]
  beta_3_pval[i] = data.frame(summary(sim_model)["coefficients"])["x3", 4]
  beta_5_pval[i] = data.frame(summary(sim_model)["coefficients"])["x5", 4]
}
```


**(d)** Based on the known values of $X$, what is the true distribution of $\hat{\beta}_1$?

**Answer:**

The true distribution of $\beta_1$ will be:

\[
\hat \beta \sim N(\beta, \sigma^2 (X^TX)^{-1})
\]

\[
\hat \beta_1 \sim N(\beta_1, \sigma ^2 C_{jj})
\]
where $\sigma ^2 = 25$, and $C_{11} = `r C[2,2]`$, so

\[
\hat \beta_1 \sim N(-0.75, `r 25 * C[2,2]`)
\]

**(e)** Calculate the mean and variance of `beta_hat_1`. Are they close to what we would expect? Plot a histogram of `beta_hat_1`. Add a curve for the true distribution of $\hat{\beta}_1$. Does the curve seem to match the histogram?

**Answer:**

```{r}
mean(beta_hat_1)
sd(beta_hat_1)
sd(beta_hat_1)^2

hist(beta_hat_1, prob = TRUE, breaks = 20,
     xlab = expression(hat(beta)[1]), border = "dodgerblue")
x = seq(-2.3, 0.6, by = 0.05)
curve(dnorm(x, mean = -0.75, sd = sqrt(25 * C[2,2])),
      col = "darkorange", add = TRUE, lwd = 3)
```


**(f)** What proportion of the p-values stored in `beta_3_pval` is less than 0.10? Is this what you would expect?

**Answer:**

```{r}
length(beta_3_pval[beta_3_pval < 0.1]) / length(beta_3_pval)
```
There is about **0.1004** of the simulations gives a p-value smaller than 0.1. This is as expected, because the true value of $\beta_3$ is 0. So, we would expect the null hypothesis where $\hat \beta_3 = 0$ gives us high p-values.

**(g)** What proportion of the p-values stored in `beta_5_pval` is less than 0.01? Is this what you would expect?

**Answer:**

```{r}
length(beta_5_pval[beta_5_pval < 0.1]) / length(beta_5_pval)
```

This is expected, because the true $\beta_5 = 2$, the cases where the null hypothesis holds should be rare, and we would expect to reject the null hypothesis for most of the time. As a result, `r length(beta_5_pval[beta_5_pval < 0.1]) / length(beta_5_pval)` of the simulations give us a p-value smaller than 0.1

