---
title: "Week 4 practice quiz"
author: "Haixu Leng"
date: "6/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Q1:** Consider a random variable $X$ that has an $F$ distribution with 33 and 55 degrees of freedom. Calculate $P[X > 2.7]$.

**A:**

```{r}
pf(2.7, 3, 5, lower.tail = FALSE)
```

**Q2:** For Questions 2-5, use the built-in longley dataset in R. Fit a multiple linear regression model with Employed as the response. Use three predictors: GNP, Population, and Armed.Forces. Specifically,

\[
Y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + \epsilon
\]

where

-$x_1$ is GNP

-$x_2$ is Population

-$x_3$ is Armed.Forces

**A:**

```{r}
longley_model = lm(Employed ~ GNP + Population + Armed.Forces, data = longley)

confint(longley_model, parm = "GNP", level = 0.9)
summary(longley_model)$coef["Population", "Std. Error"]
summary(longley_model)$coef["Armed.Forces", "Pr(>|t|)"]
summary(longley_model)$fstatistic[[1]]
```

