---
title: "week3 quiz"
author: "Haixu Leng"
date: "6/5/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1** Consider a random variable XX that has a t distribution with 5 degrees of freedom. Calculate P[|X| > 2.1].

```{r}
pt(2.1, 5, lower.tail = FALSE) * 2
```

**2** Calculate the critical value used for a 90% confidence interval about the slope parameter of a simple linear regression model that is fit to 10 observations. (Your answer should be a positive value.)  

```{r}
qt(0.95, (10 - 2))
```

**3** Consider the true simple linear regression model, 
$Y_i = 5 + 4x_i + \epsilon_i, \epsilon_i\sim N(0, \sigma^2=4), i=1,2,...20$
Given $S_{xx} = 1.5$ calculate the probability of observing data according to this model, fitting the SLR model, and obtaining an estimate of the slope parameter greater than 4.2. In other words, calculate $P[\hat \beta_1 > 4.2]$

```{r}
pnorm(0.2, 0, 2/sqrt(1.5), lower.tail = FALSE)
```

**4** Suppose we would like to predict the duration of an eruption of the Old Faithful geyser in Yellowstone National Park based on the waiting time before an eruption. Fit a simple linear model in R that accomplishes this task.

What is the value of $SE[\hat \beta_1]$?

```{r}
faithul_model = lm(eruptions~waiting, data = faithful)
summary(faithul_model)
summary(faithul_model)$coefficients[1,3]
summary(faithul_model)$coefficients[2,3]
summary(faithul_model)$coefficients[2,2]
```

**8** Calculate a 90% confidence interval for $\beta_0$. Report the upper bound of this interval.

```{r}
qt(0.95, length(faithful$waiting) - 2) * 0.160143 + (-1.874016)
```

**9** Calculate a 95% confidence interval for $\beta_1$. Report the length of the margin of this interval.

```{r}
qt(0.975, length(faithful$waiting) - 2) * summary(faithul_model)$coefficients[2,2]
(output = confint(faithul_model, level = 0.95))
(output[2,2] - output[2,1]) / 2
```

**10** Create a 90% confidence interval for the mean eruption duration for a waiting time of 81 minutes. Report the lower bound of this interval.  

```{r}
predict(faithul_model, newdata = data.frame(waiting = c(81)), interval = c("confidence"), level = 0.9)
```

**11** Create a 99% prediction interval for a new observation's eruption duration for a waiting time of 72 minutes. Report the upper bound of this interval.  

```{r}
predict(faithul_model, newdata = data.frame(waiting = c(72)), interval = c("prediction"), level = 0.99)
```

**12** Consider a 90% confidence interval for the mean response and a 90% prediction interval, both at the same xx value. Which interval is narrower?  

```{r}
set.seed(100)
x = runif(1, min(faithful$waiting), max(faithful$waiting))
predict(faithul_model, newdata = data.frame(waiting = c(x)), interval = c("confidence"), level = 0.9)
predict(faithul_model, newdata = data.frame(waiting = c(x)), interval = c("prediction"), level = 0.9)
```

**13** Fail to reject the null hypothesis



