---
title: "week3 practice quiz"
author: "Haixu Leng"
date: "6/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.  Consider a random variable X that has a t distribution with 7 degrees of freedom. Calculate P[X > 1.3].

```{r}
pt(1.3, 7, lower.tail = FALSE)
```

## 2. Consider a random variable Y that has a tt distribution with 9 degrees of freedom. Find c such that P[X > c] = 0.025

```{r}
qt(0.025, 9, lower.tail = FALSE)
```

## 3. For this exercise, use the built-in trees dataset in R. Fit a simple linear regression model with Girth as the response and Height as the predictor. What is the p-value for testing H_0: beta_1 = 0 vs H_1: beta_1 != 0?

```{r}
tree_model = lm(Girth~Height, data = trees)
summary(tree_model)$coefficients[2,4]
```

## 4.  Continue using the SLR model you fit in Question 3. What is the length of a 90% confidence interval for \beta_1?

```{r}
output = confint(tree_model, level = 0.9)
output[2,2] - output[2,1]
```

## 5. Continue using the SLR model you fit in Question 3. Calculate a 95% confidence interval for the mean tree girth of a tree that is 79 feet tall. 

```{r}
predict(tree_model, newdata = data.frame(Height = c(79)), interval = c("confidence"), level = 0.95)
```

