---
title: "Week 8 - Homework"
author: "STAT 420, Summer 2021, haixul2"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---


***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.alin = "center")
```

## Exercise 1 (Writing Functions)

**(a)** Write a function named `diagnostics` that takes as input the arguments:

- `model`, an object of class `lm()`, that is a model fit via `lm()`
- `pcol`, for controlling point colors in plots, with a default value of `grey`
- `lcol`, for controlling line colors in plots, with a default value of `dodgerblue`
- `alpha`, the significance level of any test that will be performed inside the function, with a default value of `0.05`
- `plotit`, a logical value for controlling display of plots with default value `TRUE`
- `testit`, a logical value for controlling outputting the results of tests with default value `TRUE`

The function should output:

- A list with two elements when `testit` is `TRUE`:
    - `p_val`, the p-value for the Shapiro-Wilk test for assessing normality
    - `decision`, the decision made when performing the Shapiro-Wilk test using the `alpha` value input to the function. "Reject" if the null hypothesis is rejected, otherwise "Fail to Reject."
- Two plots, side-by-side, when `plotit` is `TRUE`:
    - A fitted versus residuals plot that adds a horizontal line at $y = 0$, and labels the $x$-axis "Fitted" and the $y$-axis "Residuals." The points and line should be colored according to the input arguments. Give the plot a title. 
    - A Normal Q-Q plot of the residuals that adds the appropriate line using `qqline()`. The points and line should be colored according to the input arguments. Be sure the plot has a title. 

Consider using this function to help with the remainder of the assignment as well.

**Answer**

```{r}
# define the function
diagnostics = function(model, pcol = "grey", lcol = "dodgerblue", alpha = 0.05, plotit = TRUE, testit = TRUE){
  if (plotit){
    par(mfrow = c(1, 2))
    plot(fitted(model), resid(model), col = pcol, pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
    abline(h = 0, col = lcol, lwd = 2)
    
    qqnorm(resid(model), main = "Normal Q-Q Plot", col = pcol)
    qqline(resid(model), col = lcol, lwd = 2)
  }
  
  if (testit){
    p_val = as.numeric(shapiro.test(resid(model))["p.value"])
    l = list(p_val, ifelse(p_val < alpha, "Reject", "Fail to Reject"))
    names(l) = c("p_val", "decision")
    l
  }
}
```


**(b)** Run the following code.

```{r}
set.seed(40)

data_1 = data.frame(x = runif(n = 30, min = 0, max = 10),
                    y = rep(x = 0, times = 30))
data_1$y = with(data_1, 2 + 1 * x + rexp(n = 30))
fit_1 = lm(y ~ x, data = data_1)

data_2 = data.frame(x = runif(n = 20, min = 0, max = 10),
                    y = rep(x = 0, times = 20))
data_2$y = with(data_2, 5 + 2 * x + rnorm(n = 20))
fit_2 = lm(y ~ x, data = data_2)

data_3 = data.frame(x = runif(n = 40, min = 0, max = 10),
                    y = rep(x = 0, times = 40))
data_3$y = with(data_3, 2 + 1 * x + rnorm(n = 40, sd = x))
fit_3 = lm(y ~ x, data = data_3)
```

```{r, eval = TRUE}
diagnostics(fit_1, plotit = FALSE)$p_val
diagnostics(fit_2, plotit = FALSE)$decision
diagnostics(fit_1, testit = FALSE, pcol = "black", lcol = "black")
diagnostics(fit_2, testit = FALSE, pcol = "grey", lcol = "green")
diagnostics(fit_3)
```

***

## Exercise 2 (Prostate Cancer Data)

For this exercise, we will use the `prostate` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?prostate` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit an additive multiple regression model with `lpsa` as the response and the remaining variables in the `prostate` dataset as predictors. Report the $R^2$ value for this model.

**Answer**

```{r}
model = lm(lpsa ~ ., data = prostate)
summary(model)["r.squared"]
```

$R^2 =$ `r summary(model)["r.squared"]`.


**(b)** Check the constant variance assumption for this model. Do you feel it has been violated? Justify your answer.

**Answer**

```{r}
library(lmtest)
bptest(model)
diagnostics(model, testit = FALSE)
```

**No, it is not violated.** By looking at the fitted vs residuals plot, we find that the variance is roughly constant across different fitted values. In addition, the P value from the Breusch-Pagan test is large, so we fail to reject the constant variance hypothesis.


**(c)** Check the normality assumption for this model. Do you feel it has been violated? Justify your answer.

**Answer**

```{r}
diagnostics(model)
```

**No, it is not violated.** From the Shapiro test we did, the P value is large, so we fail to reject the normality hypothesis.

**(d)** Check for any high leverage observations. Report any observations you determine to have high leverage.

**Answer**

```{r}
leverage = hatvalues(model)
threshold = 2 * mean(leverage)
high_leverage_index = leverage > threshold
prostate[high_leverage_index, ]

plot(fitted(model), resid(model), col = "grey", pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
points(fitted(model)[high_leverage_index], resid(model)[high_leverage_index], pch = 20, col = "red", cex = 2)
abline(h = 0, col = "dodgerblue", lwd = 2)
legend("topleft", legend = c("High leverage"), col = c("red"), pch = c(20), cex = c(1))
```

The points with a leverage larger than twice the mean are listed above.

**(e)** Check for any influential observations. Report any observations you determine to be influential.

**Answer**

```{r}
influence_index = cooks.distance(model) > 4 / length(cooks.distance(model))
prostate[influence_index,]

plot(fitted(model), resid(model), col = "grey", pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
points(fitted(model)[influence_index], resid(model)[influence_index], pch = 20, col = "red", cex = 2)
abline(h = 0, col = "dodgerblue", lwd = 2)
legend("topleft", legend = c("High influence"), col = c("red"), pch = c(20), cex = c(1))
```

The points with a influence greater than $4/n$ are listed above.

**(f)** Refit the additive multiple regression model without any points you identified as influential. Compare the coefficients of this fitted model to the previously fitted model.

**Answer**

```{r}
new_data = prostate[!influence_index, ]
new_model = lm(lpsa ~ ., data = new_data)
coef(model)
coef(new_model)
```
By comparing the coefficients, we have the following observations:

  - Intercept: changes from a positive number to a negative number
  
  - lcavol, lweight, age, lbph, svi, lcp, pgg45: the change is relatively small
  
  - gleason: this coefficient increases more than 100 %

**(g)** Create a data frame that stores the observations that were "removed" because they were influential. Use the two models you have fit to make predictions with these observations. Comment on the difference between these two sets of predictions.

**Answer**

```{r}
removed = prostate[influence_index, ]
predict(model, removed)
predict(new_model, removed)
# real value
plot(predict(model, removed), predict(new_model, removed),
     xlab = "full model",
     ylab = "model without influential points",
     main = "prediction comparison")
abline(0, 1, col = "orange")
legend("topleft", legend = "y = x", col = "orange", lwd = "1")
removed$lpsa
```
Neither of the two models match the real value. Actually, the predictions from the two models are quite close to each other.

***

## Exercise 3 (Why Bother?)

**Why** do we care about violations of assumptions? One key reason is that the distributions of the parameter esimators that we have used are all reliant on these assumptions. When the assumptions are violated, the distributional results are not correct, so our tests are garbage. **Garbage In, Garbage Out!**

Consider the following setup that we will use for the remainder of the exercise. We choose a sample size of 50.

```{r}
n = 50
set.seed(420)
x_1 = runif(n, 0, 5)
x_2 = runif(n, -2, 2)
```

Consider the model,

\[
Y = 4 + 1 x_1 + 0 x_2 + \epsilon.
\]

That is,

- $\beta_0$ = 4
- $\beta_1$ = 1
- $\beta_2$ = 0

We now simulate `y_1` in a manner that does **not** violate any assumptions, which we will verify. In this case $\epsilon \sim N(0, 1).$

```{r}
set.seed(83)
library(lmtest)
y_1 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = 1)
fit_1 = lm(y_1 ~ x_1 + x_2)
bptest(fit_1)
```

Then, we simulate `y_2` in a manner that **does** violate assumptions, which we again verify. In this case $\epsilon \sim N(0, \sigma = |x_2|).$

```{r}
set.seed(83)
y_2 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = abs(x_2))
fit_2 = lm(y_2 ~ x_1 + x_2)
bptest(fit_2)
```

**(a)** Use the following code after changing `birthday` to your birthday.

```{r}
num_sims = 2500
p_val_1 = rep(0, num_sims)
p_val_2 = rep(0, num_sims)
birthday = 19900210
set.seed(birthday)
```

Repeat the above process of generating `y_1` and `y_2` as defined above, and fit models with each as the response `2500` times. Each time, store the p-value for testing,

\[
\beta_2 = 0,
\]

using both models, in the appropriate variables defined above. (You do not need to use a data frame as we have in the past. Although, feel free to modify the code to instead use a data frame.)

**Answer**

```{r}
for (i in 1:num_sims){
  y_1 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = 1)
  fit_1 = lm(y_1 ~ x_1 + x_2)
  p_val_1[i] = summary(fit_1)$coefficients[3,4]
  
  y_2 = 4 + 1 * x_1 + 0 * x_2 + rnorm(n = n, mean = 0, sd = abs(x_2))
  fit_2 = lm(y_2 ~ x_1 + x_2)
  p_val_2[i] = summary(fit_2)$coefficients[3,4]
}
```


**(b)** What proportion of the `p_val_1` values is less than 0.01? Less than 0.05? Less than 0.10? What proportion of the `p_val_2` values is less than 0.01? Less than 0.05? Less than 0.10? Arrange your results in a table. Briefly explain these results.

**Answer**

```{r}
table = data.frame(c("No Violation", "Violation"), 
                   c(mean(p_val_1[p_val_1 < 0.01]), mean(p_val_2[p_val_2 < 0.01])),
                   c(mean(p_val_1[p_val_1 < 0.05]), mean(p_val_2[p_val_2 < 0.05])),
                   c(mean(p_val_1[p_val_1 < 0.1]), mean(p_val_2[p_val_2 < 0.1])))
names(table) = c("", "p<0.01", "p<0.05", "p<0.1" )
knitr::kable(table)

par(mfrow = c(1, 2))
hist(p_val_1)
hist(p_val_2)
par(mfrow = c(1, 1))
plot(c(0.01, 0.05, 0.1), c(mean(p_val_1[p_val_1 < 0.01]),mean(p_val_1[p_val_1 < 0.05]), mean(p_val_1[p_val_1 < 0.1])),
     pch = 20,
     cex = 2,
     xlab = "P value",
     ylab = "Proportion",
     main = "Comparison")

points(c(0.01, 0.05, 0.1), c(mean(p_val_2[p_val_2 < 0.01]),mean(p_val_2[p_val_2 < 0.05]), mean(p_val_2[p_val_2 < 0.1])),
     col = "red", pch = 20, cex = 2)
legend("topleft", legend = c("No Violation", "Violation"),
       col = c("black", "red"),
       pch = c(20, 20))
```

From the value we can find that the proportion where the p-value is less than x (x = 0.01, 0.05, 0.1) is different with respect to different x. For example, in the case of no violation, the p-value follows a uniform distribution. So, when x increases, the proportion that p is less than x **increases linearly**. However, for the case with violation, the p distribution is left skewed. So, when x increases, the proportion that p is less than x increases **slower (non-linear)** than the case with no violation.


***

## Exercise 4 (Corrosion Data)

For this exercise, we will use the `corrosion` data, which can be found in the `faraway` package. After loading the `faraway` package, use `?corrosion` to learn about this dataset.

```{r, message = FALSE, warning = FALSE}
library(faraway)
```

**(a)** Fit a simple linear regression with `loss` as the response and `Fe` as the predictor. Plot a scatterplot and add the fitted line. Check the assumptions of this model.

**Answer**

```{r}
model = lm(loss ~ Fe, data = corrosion)
plot(corrosion$Fe, corrosion$loss,
     xlab = "Fe",
     ylab = "loss",
     main = "Loss vs Fe")
abline(model, col = "orange")
bptest(model)
diagnostics(model)
```

We checked assumptions by doing Breusch-Pagan test and the Shapiro test. In both tests, the p-value is fairly large, so we fail to reject the null hypothesis. No violation is found.

**(b)** Fit higher order polynomial models of degree 2, 3, and 4. For each, plot a fitted versus residuals plot and comment on the constant variance assumption. Based on those plots, which of these three models do you think are acceptable? Use a statistical test(s) to compare the models you just chose. Based on the test, which is preferred? Check the normality assumption of this model. Identify any influential observations of this model.

**Answer**

```{r}
par(mfrow = c(1, 3))
model2 = lm(loss ~ poly(Fe, degree = 2), data = corrosion)
plot(fitted(model2), resid(model2), col = "grey", pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals, degree 2")
abline(h = 0, col = "dodgerblue", lwd = 2)

model3 = lm(loss ~ poly(Fe, degree = 3), data = corrosion)
plot(fitted(model3), resid(model3), col = "grey", pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals, degree 3")
abline(h = 0, col = "dodgerblue", lwd = 2)

model4 = lm(loss ~ poly(Fe, degree = 4), data = corrosion)
plot(fitted(model4), resid(model4), col = "grey", pch = 20,
         xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals, degree 4")
abline(h = 0, col = "dodgerblue", lwd = 2)
```

Based on the fitted versus residuals plots, I think the model with **degree = 3** is acceptable, because the other two has large residuals on one side of the x axis.

```{r}
bptest(model2)
bptest(model3)
bptest(model4)
```

Based on the BP test, all three models have a large p value. But the model with **degree = 2** has the largest p value, so it is preferred. The one I picked (degree = 3) based on the plots actually have the least p value, which is interesting.

```{r}
diagnostics(model2)
diagnostics(model3)
```

If I run diagnostic test on models with degree 2 and 3, the model with **degree = 3** looks better. For **model with degree = 3**, **no influential points are found**.

```{r}
influence_index = cooks.distance(model3) > 4 / length(cooks.distance(model3))
corrosion[influence_index,]
```


***

## Exercise 5 (Diamonds)

The data set `diamonds` from the `ggplot2` package contains prices and characteristics of 54,000 diamonds. For this exercise, use `price` as the response variable $y$, and `carat` as the predictor $x$. Use `?diamonds` to learn more.

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
```

**(a)** Fit a linear model with `price` as the response variable $y$, and `carat` as the predictor $x$. Return the summary information of this model.

**Answer**

```{r}
model = lm(price ~ carat, data = diamonds)
summary(model)
```


**(b)** Plot a scatterplot of price versus carat and add the line for the fitted model in part **(a)**. Using a fitted versus residuals plot and/or a Q-Q plot, comment on the diagnostics. 

**Answer**

```{r}
plot(diamonds$carat, diamonds$price,
     xlab = "carat",
     ylab = "price",
     main = "Price vs Carat")
abline(model, col = "orange")

diagnostics(model, testit = FALSE)
```

By checking the fitted vs residual plot, we find that the equal variance **does not hold**. At low prices the variance seems to be greater than the variance at high prices.

By checking the Q-Q plot, we find that the model and data agree at the center section. However, at two ends **the result deviates from the normal distribution assumption**.

**(c)** Seeing as the price stretches over several orders of magnitude, it seems reasonable to try a log transformation of the response. Fit a model with a logged response, plot a scatterplot of log-price versus carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

```{r}
qplot(price, data = diamonds, bins = 30)
```


**Answer**

```{r}
model_log = lm(log(price) ~ carat, data = diamonds)
plot(diamonds$carat, log(diamonds$price),
     xlab = "carat",
     ylab = "log(price)",
     main = "log(Price) vs Carat")
abline(model_log, col = "orange")

diagnostics(model_log, testit = FALSE)
```

By checking the fitted vs residual plot, we find that the equal variance **does not hold**. At low prices the variance seems to be greater than the variance at high prices.

By checking the Q-Q plot, we find that the model and data agree at the center section and the upper end. However, at the lower end **the result deviates from the normal distribution assumption**.


**(d)** Try adding log transformation of the predictor. Fit a model with a logged response and logged predictor, plot a scatterplot of log-price versus log-carat and add the line for the fitted model, then use a fitted versus residuals plot and/or a Q-Q plot to comment on the diagnostics of the model.

**Answer**

```{r}
model_log_log = lm(log(price) ~ log(carat), data = diamonds)
plot(log(diamonds$carat), log(diamonds$price),
     xlab = "log(carat)",
     ylab = "log(price)",
     main = "log(Price) vs log(Carat)")
abline(model_log_log, col = "orange")

diagnostics(model_log_log, testit = FALSE)
```

By checking the fitted vs residual plot, we find that the equal variance **does hold**. At high prices the variance seems to be slightly greater than the variance at low prices.

By checking the Q-Q plot, we find that the model and data **agree at most places**. However, at the both ends the result deviates a bit from the normal distribution assumption.

**(e)** Use the model from part **(d)** to predict the price (in dollars) of a 3-carat diamond. Construct a 99% prediction interval for the price (in dollars).

**Answer**

```{r}
exp(predict(model_log_log, data.frame(carat = 3), interval = "prediction", level = 0.99))
```
The interval is listed above. As a quality control step, I googled price of a 3 carat diamond, and I found the price to be around 20 k. So, our prediction is very close.

